---
title: "Covid-19 Nursing home Mortality rate"
author: "Sina Mokhtar"
date: '2022-06-21'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(usdata)
```


## load nursery home datasets

```{r}
load( "Data/nurse_df.csv")
```


## Top 5 state with cumulative mortality rate

```{r}
# summerize data to state level and filter top 5 state with cumulative mortality rate
# (after "2021-06-30")
# Note: mortality rate = death / occupied bed
nurse_state_cum_top5_list = nurse_df %>%
    drop_na(state) %>%
    group_by(date, state) %>%
    summarize(deaths = sum(deaths, na.rm = TRUE),
              occupied_bed = sum(occupied_bed, na.rm = TRUE)) %>%
    mutate(mortalityRate = (deaths / occupied_bed) ) %>%
    mutate(state_ab = state2abbr(state)) %>%
    dplyr::filter(date >= "2021-06-30") %>%
    group_by(state) %>%
    mutate(mortalityRate_cum = cumsum(mortalityRate)) %>%
    group_by(state) %>%
    dplyr::arrange(desc(mortalityRate_cum)) %>%
    dplyr::select(state) %>%
    distinct() %>%
    head(5)
```
    
## Compute county level mortality rate and vaccine rate

```{r}
# summarize data to county level and filter by date and top 5 state
# Note: mortality rate = death / occupied bed
# Note: accupancyRate = occupied_bed / total_bed
nurse_county = nurse_df %>%
    drop_na(state, county)  %>%
    filter(date >= "2021-06-30") %>%
    filter(state %in% nurse_state_cum_top5_list$state) %>%
    group_by(date, state, county) %>%
    mutate(nurse_shortage = dplyr::case_when(nurse_shortage == "Y" ~ 1,
                                             nurse_shortage == "N" ~ 0))  %>%
    summarize(deaths = sum(deaths, na.rm = TRUE),
              occupied_bed = sum(occupied_bed, na.rm = TRUE),
              nurse_shortage = mean(nurse_shortage, na.rm = TRUE),
              total_bed = sum(total_bed, na.rm = TRUE),
              resident_vaccineRate = mean(resident_vaccineRate, na.rm = TRUE),
              personnel_vaccineRate = mean(personnel_vaccineRate, na.rm = TRUE),
              resident_vaccineRate_boost = mean(resident_vaccineRate_boost, na.rm = TRUE),
              personnel_vaccineRate_boost = mean(personnel_vaccineRate_boost, na.rm = TRUE))%>% 
    ungroup() %>%
    mutate(mortalityRate = (deaths / occupied_bed) * 100,
           occupancyRate = (occupied_bed / total_bed) * 100,
           state_abb = state2abbr(state),
           occupancyRate = if_else(occupancyRate < 100, occupancyRate, NULL)) %>%
    drop_na(mortalityRate) %>%
    mutate(mortalityRate = ifelse(mortalityRate == Inf, NA, mortalityRate))
```

\newpage

## Linear models for mortalityRate VS nurse_shortage

```{r}
fit = lm(mortalityRate ~ nurse_shortage,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_nurse_shortage = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = nurse_shortage)) +
  labs(title = "mortalityRate VS nurse_shortage") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_nurse_shortage 

```
   
\newpage   
   
## Linear models for mortalityRate VS resident_vaccineRate

```{r}
fit = lm(mortalityRate ~ resident_vaccineRate,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_resident_vaccineRate = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = resident_vaccineRate)) +
  labs(title = "mortalityRate VS resident_vaccineRate") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_resident_vaccineRate
```
\newpage

## Linear models for mortalityRate VS personnel_vaccineRate

```{r}
fit = lm(mortalityRate ~ personnel_vaccineRate,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_personnel_vaccineRate = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = personnel_vaccineRate)) +
  labs(title = "mortalityRate VS personnel_vaccineRate") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_personnel_vaccineRate 
```
\newpage

## Linear models for mortalityRate VS resident_vaccineRate_boost

```{r}
fit = lm(mortalityRate ~ resident_vaccineRate_boost,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_resident_vaccineRate_boost = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = resident_vaccineRate_boost)) +
  labs(title = "mortalityRate VS resident_vaccineRate_boost") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_resident_vaccineRate_boost 
```
   
\newpage   
    
## Linear models for mortalityRate VS personnel_vaccineRate_boost

```{r}
fit = lm(mortalityRate ~ personnel_vaccineRate_boost,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_personnel_vaccineRate_boost = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = personnel_vaccineRate_boost)) +
  labs(title = "mortalityRate VS personnel_vaccineRate_boost") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_personnel_vaccineRate_boost
``` 
\newpage

## Linear models for mortalityRate VS occupancyRate

```{r}
fit = lm(mortalityRate ~ occupancyRate,
         data = nurse_county)

summary(fit)
#plot the model
fig_mortalityRate_occupancyRate = ggplot(data = nurse_county,
                                aes(y = mortalityRate,
                                    x = occupancyRate)) +
  labs(title = "mortalityRate VS occupancyRate") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_mortalityRate_occupancyRate
```   


\newpage   
    
## Time series of the nurse_shortage
```{r}
fig_time_series_county_nurse_shortage = ggplot(data = nurse_county,
                         aes(x = date,
                             y = nurse_shortage))  +
  labs(title = "nurse_shortage in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
ggsave("Result/fig_time_series_county_nurse_shortage.jpg",
       fig_time_series_county_nurse_shortage,
       height=4,
       width=8,
       scale=1.65)

```


\newpage   
    
## Time series of the resident_vaccineRate
```{r}
fig_time_series_county_resident_vaccineRate = ggplot(data = nurse_county,
                         aes(x = date,
                             y = resident_vaccineRate))  +
  labs(title = "resident_vaccineRate in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_resident_vaccineRate
ggsave("Result/fig_time_series_county_resident_vaccineRate.jpg",
       fig_time_series_county_resident_vaccineRate,
       height=4,
       width=8,
       scale=1.65)

```


\newpage   
    
## Time series of the personnel_vaccineRate
```{r}
fig_time_series_county_personnel_vaccineRate = ggplot(data = nurse_county,
                         aes(x = date,
                             y = personnel_vaccineRate))  +
  labs(title = "personnel_vaccineRate in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_personnel_vaccineRate
ggsave("Result/fig_time_series_county_personnel_vaccineRate.jpg",
       fig_time_series_county_personnel_vaccineRate,
       height=4,
       width=8,
       scale=1.65)

```



\newpage   
    
## Time series of the resident_vaccineRate_boost
```{r}
fig_time_series_county_resident_vaccineRate_boost = ggplot(data = nurse_county,
                         aes(x = date,
                             y = resident_vaccineRate_boost))  +
  labs(title = "resident_vaccineRate_boost in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_resident_vaccineRate_boost
ggsave("Result/fig_time_series_county_resident_vaccineRate_boost.jpg",
       fig_time_series_county_resident_vaccineRate_boost,
       height=4,
       width=8,
       scale=1.65)

```


\newpage   
    
## Time series of the personnel_vaccineRate_boost
```{r}
fig_time_series_county_personnel_vaccineRate_boost = ggplot(data = nurse_county,
                         aes(x = date,
                             y = personnel_vaccineRate_boost))  +
  labs(title = "personnel_vaccineRate_boost in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_personnel_vaccineRate_boost
ggsave("Result/fig_time_series_county_personnel_vaccineRate_boost.jpg",
       fig_time_series_county_personnel_vaccineRate_boost,
       height=4,
       width=8,
       scale=1.65)

```

\newpage   
    
## Time series of the occupancyRate
```{r}
fig_time_series_county_occupancyRate = ggplot(data = nurse_county,
                         aes(x = date,
                             y = occupancyRate))  +
  labs(title = "occupancyRate in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_occupancyRate
ggsave("Result/fig_time_series_county_occupancyRate.jpg",
       fig_time_series_county_occupancyRate,
       height=4,
       width=8,
       scale=1.65)

```


\newpage   
    
## Time series of the mortalityRate
```{r}
fig_time_series_county_mortalityRate = ggplot(data = nurse_county,
                         aes(x = date,
                             y = mortalityRate))  +
  labs(title = "mortalityRate in county level from July 2021 ") +
  geom_point() +
  geom_smooth(method='lm',
              formula= y~x,
              se = FALSE)
fig_time_series_county_mortalityRate
ggsave("Result/fig_time_series_county_mortalityRate.jpg",
       fig_time_series_county_mortalityRate,
       height=4,
       width=8,
       scale=1.65)

```



